---
title: "intro_dplyr_en"
output: html_document
---


# Init
## Create an new R project.

## load here the packages {tidyverse} and {prenoms} you need. You will add other required package in this chunk later.

_ the {prenoms} package isn't on CRAN, if missing you can install it with : `remotes::install_github( "ThinkR-open/prenoms" )`

```{r}






```

# Import
## Using `data(prenoms)` preload the `prenoms` example prenoms from the {prenoms} packages. This is a function that will load the database without using the assignment arrow `<-`. the `data` function must not be used in any other case: only to import example dataset from a package.


```{r}

```

## What kind of object is prenoms

```{r}

```



## Use the `5-functions-you-have-to-always-run` on the prenoms object

```{r}





```

_sometimes the `View` function does not allow you to compile your notebook. Feel free to comment this line._



## Using `glimpse`, have a look to `prenoms` structure.


```{r}

```


_In this Rmd you dont have to assign your results to an object, you dont have to replace/edit/transform the dataset `prenoms`. You only have to show the output in the rendered document._


# Sorting the dataset 

## Sort the dataset by year

```{r}


```

## Sort the dataset by decreasing number of occurrences

```{r}


```

## Sort the dataset by name in alphabetical order and by decreasing proportion

```{r}


```


# Filtrer the dataset


## Display only available information on male names

```{r}


```


## Display only the lines corresponding to girls born in 2011


```{r}




```


## Keep in the database only information on the first names "Vincent", "Diane", "Colin", "Sébastien" and "Cervan".

```{r}


```

# Select

## Keep only the `name`, `year` and `n` columns .

```{r}


```

## Keep all variables except `sex` et `prop`

```{r}


```


## Keep only variables that contain an "e" in their name. _(The name of the variable (= column header), not the first birth names inside variable "name" of this dataset !!)_


```{r}


```


## Keep only variables whose names end with an "e"

```{r}


```



# Transformation 

## Multiply by 100 and do a relevant rounding on the `prop` variable

```{r}



```


## Currently the `prop` column is calculated by sex and year, add another proportion column, calculated globally over the entire dataset. (No, this new variable does not make sense at all, so what ?)



```{r}

```


# Resume 


## How many babies were born in France between 1900 and 2017?


```{r}


```

## How many births were there in France in 1925?


```{r}



```

## How many births per year have there been in France?


```{r}




```

# Explorer

## How many different first names exist in the `prenoms` database?

```{r}



```


## How many times has your first name been given in your department in your year of birth? (foreigner ? Use mine. My name is Vincent. I was born in 1986 in the department 76 :) )

```{r}


```



## Find the most popular female name in Paris in 1950.

```{r}









```



## What are the most common names given, by gender, year and department?

```{r}



```



## What is the name that has been given most in absolute value all together?

```{r}




```



## What is the most common boy's name given, all combined?

```{r}





```

## Build a table that track the popularity of your first name over time ( You can use any name you want ) _(Each year, the number of times your first name has been given)_

```{r}




```

## Build a database called `france` that removes the department scale (you must aggregate the departments by year and gender)

- Do not forget to add the variable "proportion" to `france`. _(this variable must make sense... think carefully about its meaning)_

```{r}











```

## In the `france` database, using `case_when` in the sex column, replace "M" with "Male" and "F" with "Female".

```{r}







```



## Using the `france` database, only print information about names ending with the letter `z`.

_You may want to look at the {stringr} package to do this_

```{r}


```


## Make a database with 3 columns: `year`, `most given male name`, `most given female name`, each line contains a year and 2 names.

```{r}








```

